---
title: 'Tidy Tuesday: 9/15/2020 - US Spending on Kids'
author: "Ashlyn Johnson"
date: "9/15/2020"
output: github_document
---

## Packages
```{r packages, message = FALSE}
library(tidyverse)
library(tidytuesdayR)
library(viridis)
```

## Data   
This week's data comes from the Urban Institute and we will use Joshua Rosenberg's package, `tidykids` to access the data. The dataset provides information on public spending on children from 1997 - 2016. The `tidykids` package contains a cleaned and tidy version of the data. 

```{r importing data, results = 'hide', message = FALSE, error = FALSE, warning = FALSE}
dat <- tidytuesdayR::tt_load(2020, week = 38)
kids <- dat$kids
```

**inf_adj** is the value adjusted for inflation and **inf_adj_perchild** is the value adjusted for inflation per child.   

### Initial Plots to Check Out the Data

```{r fig1, fig.width = 12, fig.height = 20, tidy = TRUE}
ggplot() + 
  geom_area(data = kids %>% 
              filter(variable == "PK12ed"), 
            aes(x = year, y = inf_adj_perchild, fill = state), 
            show.legend = FALSE) + 
  scale_fill_viridis(discrete = TRUE) + 
  theme_classic() + 
  labs(title = "Public Spending on Elementary & Secondary Schools Per Child by State", subtitle = "in $1000s", xlab = "Year", ylab = "$1000s of Dollars per Child, Adjusted for Inflation") +
  facet_wrap(~state, ncol = 5)
```




