---
title: 'Tidy Tuesday: 9/15/2020 - US Spending on Kids'
author: "Ashlyn Johnson"
date: "9/15/2020"
output: github_document
---

## Packages
```{r packages, message = FALSE}
library(tidyverse)
library(tidytuesdayR)
library(viridis)
library(urbnmapr)
library(gganimate)
library(transformr)
library(ggthemes)
library(gifski)
```

## Data   
This week's data comes from the Urban Institute and we will use Joshua Rosenberg's package, `tidykids` to access the data. The dataset provides information on public spending on children from 1997 - 2016. The `tidykids` package contains a cleaned and tidy version of the data. 

```{r importing data, results = 'hide', message = FALSE, error = FALSE, warning = FALSE}
dat <- tidytuesdayR::tt_load(2020, week = 38)
kids <- dat$kids %>% 
  rename(state_name = state)
```

**inf_adj** is the value adjusted for inflation and **inf_adj_perchild** is the value adjusted for inflation per child.   

### Initial Plots to Check Out the Data

```{r fig1, fig.width = 12, fig.height = 20, tidy = TRUE}
ggplot() + 
  geom_area(data = kids %>% 
              filter(variable == "PK12ed"), 
            aes(x = year, y = inf_adj_perchild, fill = state_name), 
            show.legend = FALSE) + 
  scale_fill_viridis(discrete = TRUE) + 
  theme_classic() + 
  labs(title = "Public Spending on Elementary & Secondary Schools Per Child by State", subtitle = "in $1000s", xlab = "Year", ylab = "$1000s of Dollars per Child, Adjusted for Inflation") +
  facet_wrap(~state_name, ncol = 5)
```


Looking at this initial plot, I am curious to see this data in more concise way. Can I plot this data onto a map of the US?

### US Map Based Plots

This is a static plot for the year 2000. 

```{r fig2, fig.width = 12, fig.height = 8}
states_sf <- get_urbn_map("states", sf = TRUE)


kids_spatial_2000 <- full_join(states_sf,
                          kids %>% 
                            filter(variable == "PK12ed", year == 2000),
                          by = "state_name")

kids_spatial_2000 %>% 
ggplot() +
  geom_sf(mapping = aes(fill = inf_adj_perchild),
          color = "black", size = 0.25) +
  scale_fill_viridis(discrete = FALSE) +
   geom_sf_text(data = get_urbn_labels(map = "states", sf = TRUE), 
                aes(label = state_abbv), color = "black", 
            size = 5) + 
  labs(fill = "thousand dollars",
       title = "Public Spending on Elementary and Secondary Schools: 2000",
       subtitle = "in thousands of dollars per child, adjusted for inflation") +
  theme_map() + 
  theme(plot.title = element_text(size = 20, face = "bold"),
        plot.subtitle = element_text(size = 16),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10),
        legend.key.size = unit(20, "pt"),
        legend.position = "top",
        panel.background = element_rect(fill = 'gray', colour = 'gray'),
        plot.background = element_rect(fill = 'gray', colour = 'gray'), 
        legend.background = element_rect(fill = "gray", color = "gray"))
```


This is an animated plots to show all of the years. 

```{r animated plot}

kids_spatial <- full_join(states_sf,
                          kids %>% 
                            filter(variable == "PK12ed"),
                          by = "state_name")

plot <-kids_spatial %>% 
ggplot() +
  geom_sf(mapping = aes(fill = inf_adj_perchild),
          color = "black", size = 0.25) +
  scale_fill_viridis(discrete = FALSE) +
   geom_sf_text(data = get_urbn_labels(map = "states", sf = TRUE), 
                aes(label = state_abbv), color = "black", 
            size = 15) + 
  labs(fill = "thousand dollars",
       subtitle = "in thousands of dollars per child, adjusted for inflation") +
  theme_map() + 
  theme(plot.title = element_text(size = 50, face = "bold"),
        plot.subtitle = element_text(size = 40),
        legend.title = element_text(size = 30, face = "bold"),
        legend.text = element_text(size = 25),
        legend.key.size = unit(60, "pt"),
        legend.position = "top",
        panel.background = element_rect(fill = 'gray', colour = 'gray'),
        plot.background = element_rect(fill = 'gray', colour = 'gray'), 
        legend.background = element_rect(fill = "gray", color = "gray"))


 plot_animated <- plot +
  ggtitle("Public Spending on Elementary and Secondary Schools: {frame_time}") +
  transition_time(as.integer(year)) +
  ease_aes("linear") +
  enter_fade() +
  exit_fade()
  
 
animate(plot_animated, renderer = gifski_renderer(), width = 1000, height = 666, res = 35)
anim_save("kids_school_spending.gif")
```


